---
title: "Your Paper Title Here"
author: "Your Name(s)"
date: Date
output: pdf_document
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE, warning=FALSE}
#Use this code chunk to include libraries, and set global options.
library(tidyverse)
```

# Abstract
Write your abstract here.

# Keywords
*Keyword 1*, *Keyword 2*, *Keyword 3*, *Keyword 4*

\newpage

# Introduction / Background
Provide an introduction with background information.

```{r label="important_R_code", eval=TRUE, echo=FALSE}
# You can reference your code in the appendix (sample here).
```

# Statistical Methods

## Model
Describe the statistical model used.

## Likelihood Inference
Detail the likelihood approach.

## Bayesian Inference
Detail the Bayesian approach.


Let's say we believe 

$$
\begin{aligned}
P(\psi \geq 0.3) &= P(\frac{1-2\pi}{1-\pi} \geq 0.3)=0.1 \rightarrow P(\pi \leq \frac{7}{17}) = 0.1 \\
P(\psi \geq 0) &= P(\frac{1-2\pi}{1-\pi} \geq 0)=0.1\rightarrow P(\pi \leq \frac{1}{2}) = 0.5
\end{aligned}
$$
```{r, warning=FALSE}
library(LearnBayes)
beta.select(quantile1 = list(p = 0.1, x = 7/17),
            quantile2 = list(p = 0.5, x = 0.5))
```

That is,
$$
P(\pi | x)  \sim Beta(26.2+8 , 26.2+162)
$$





$$
\begin{aligned}
P(\Psi \leq \psi | x) &=P(\frac{1-2\pi}{1-\pi} \leq \psi|x) \quad -\infty  < \psi \leq 1 \\
&= P(\pi \geq \frac{\psi-1}{\psi-2}|x) \\
&= 1- P(\pi  \leq \frac{\psi-1}{\psi-2}|x )\\
F_{\Psi}(\psi|x) &= 1- F_{\pi}(\frac{\psi-1}{\psi-2}|x) \\
f_{\Psi}(\psi|x) &=  (1- F_{\pi}(\frac{\psi-1}{\psi-2}|x))'(\frac{\psi-1}{\psi-2})'\\
f_{\Psi}(\psi|x) &= f_{\pi}(\frac{\psi-1}{\psi-2}|x)\frac{1}{(\psi -2)^2} \\
&= \frac{(\frac{\psi-1}{\psi-2})^{33.2}(\frac{-1}{\psi-2})^{187.2}}{B(34.2, 188.2)}\frac{1}{(\psi -2)^2} \\
&= \frac{(1-\psi)^{33.2}}{B(34.2, 188.2)(2-\psi)^{222.4}} \quad  -\infty< \psi \leq 1
\end{aligned}
$$
The prior for $\Psi$ will be 
$$
f_{\Psi}(\psi) = \frac{(1-\psi)^{25.2}}{B(26.2, 26.2)(2-\psi)^{52.4}} \quad  -\infty< \psi \leq 1
$$






```{r}
ggplot()+
  geom_function(
    fun = function(x) {(1-x)^33.2/(beta(34.2, 188.2)*(2-x)^222.4)},
    xlim = c(0, 1),
    aes(color = 'Poserior'),
    lwd = 1.2
  )+
  geom_function(
    fun = function(x) {(1-x)^25.2/(beta(26.2, 26.2)*(2-x)^52.4)},
    xlim = c(0, 1),
    aes(color = 'Prior'),
    lwd = 1.2
  )+
  labs(x = expression(psi), title= paste('Posterior distribution of psi') )+
  theme_classic()

```




# Results

To perform inference 

Present your findings. 

Median of posterior distribution: 

$P(\psi \leq q|u)$ \newline 
$P(1-2\pi \leq (q|u)(1-\pi))$ \newline 

```{r}
inverse <- qbeta(0.5, 34.2, 188.2, lower.tail = F)
q <- (2*inverse - 1)/(inverse - 1 )
```

To perform inference with Bayesian methods, summary statistics of the posterior distribution must be reported. For this report, the posterior median will be reported.

To find the posterior median, 
\begin{alignat*}{2}
&P(\psi \leq q | v_{obs}) = 0.5 &  \\ 
&P(\frac{1-2\pi}{1-\pi} \leq q | v_{obs}) =  0.5 & \text{Definition of } \psi \\ 
&P(1-2\pi \leq q_v - q_v \pi) = 0.5 & \\ 
&P(1-q_v \leq 2\pi - q_v\pi = 0.5) & \\ 
&P(\frac{1-q_v}{2-q_v} \leq \pi) = 0.5 & \\ 
&1 - P(\frac{1-q_v}{2-q_v} \geq \pi) = 0.5 & \text{Definition of complement} \\ 
&\frac{1-q_v}{2-q_v} = \approx 0.1527383 & \text{Inverting probability function} \\
&(1-q_v) = (2-q_v)(0.1527383) & \\ 
&q_v = \frac{2*0.1527383-1}{0.1527383 - 1} \\ 
&q_v = 0.8197271 &
\end{alignat*}

The posterior median of Pfizer's COVID-19 efficacy, $\psi$, is 0.819271. To fully contextualize this, the upper and lower bounds of the credible interval are:
Lower Bound $= q_v = \frac{2*0.1095326-1}{0.1095326 - 1}, \quad 0.1095326$ is 0.005 percentile of posterior distribution. 

Upper Bound $= q_v = \frac{2*0.2039136-1}{0.2039136 - 1}, \quad 0.2039136$ from 99.5 percentile of posterior distribution.


```{r lower and upper bounds}
lower_q = qbeta(0.005, shape1 = 34.2, shape2 = 188.2)
upper_q = qbeta(0.995, shape1 = 34.2, shape2 = 188.2)

lowbound <- (2*lower_q-1)/(lower_q - 1)
upbound <- (2*upper_q - 1)/(upper_q - 1)
```

Our credible interval is [`r lowbound`,`r upbound`]. This means there is a 99% probability that the real value of $\psi$ lies within this interval. 

For the frequentist method, 

# Discussion / Conclusion
Discuss / conclude here.

# Bibliography

Brown, B. (2024). *Lecture Title*. Lecture slides, Course Name, University Name.

Doe, J. (2020). Title of the Paper. *Journal Name*, 12(3), 45-67.

Last, F., & Last, F. (2025). *Book Title*. Publisher.

Smith, A., & Johnson, C. (2023). *Title of the Online Article*. Retrieved from https://www.example.com.

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

## Proofs
If applicable, include detailed mathematical derivations or additional theoretical explanations.
