---
title: "Your Paper Title Here"
author: "Blake Xu, Kevin Ji, Sangey Rinchen"
date: Date
output: pdf_document
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.
library(tidyverse)
library(fastR2)
```

# Abstract
Write your abstract here.

# Keywords
*Keyword 1*, *Keyword 2*, *Keyword 3*, *Keyword 4*

\newpage

# Introduction / Background
Provide an introduction with background information.

```{r label="important_R_code", eval=TRUE, echo=FALSE}
# You can reference your code in the appendix (sample here).
```

# Statistical Methods

## Model

We define $V$ as the proportion of people who were infected with COVID-19 that received the vaccine. $n$ is the total number of test subjects that were infected (170 in our case) and $\pi_0$ is the true but unknown probability of being vaccinated and still getting infected. 

This true probability has been calculated in terms of two other probabilities, $\pi_v$ and $\pi_p$, which are the probabilities that the subject received the vaccine and a placebo respectively:  

$$\pi = \frac{\pi_v}{\pi_v + \pi_p}$$  

What we are interested in is the efficacy of the vaccine, $\psi$. This efficacy, according to the study, is computed with the following formula:  

$$\psi = \frac{1 - 2\pi}{1 - \pi}$$  

We can transform this variable in terms of $\psi$ as well (which is more helpful for our inferences):  

$$\pi = \frac{1 - \psi}{2 - \psi}$$

## Likelihood Inference
### Likelihood function and MLE

In this section, we take the likelihood approach and break down the problem through a frequentist's lens. We begin by approximating $V$, the proportion of subjects infected with COVID-19 that were vaccinated, to the binomial distribution (proof in appendix):  

$$V \approx Binom(n, \pi_0)$$

In this case, we know $n = 170$ and observe $v = 8$. In our lectures, we proved that the MLE of a binomial distribution is the proportion of successes to the number of trials. Hence,  

$$\widehat{\pi}_0^{mle} = \frac{8}{170}$$  
Then, with our formula for $\psi$,  

$$\widehat{\psi}_0^{mle} = \frac{1 - 2(\frac{8}{170})}{1 - (\frac{8}{170})} \approx 0.9506$$  

Let us verify this with the transformed likelihood function, $L^{\ast}(\psi)$:  

$$L(\pi) = \binom{n}{v}\pi^v(1-\pi)^{n-v}$$  
$$L^{\ast}(\psi) = \binom{n}{v} (\frac{1 - \psi}{2 - \psi})^v (1- \frac{1 - \psi}{2 - \psi})^{n-v}$$  
$$L^{\ast}(\psi) = \binom{n}{v} (\frac{1 - \psi}{2 - \psi})^v (\frac{1}{2 - \psi})^{n - v}$$  
$$L^{\ast}(\psi) = \binom{n}{v}\frac{(1 - \psi)^v}{(2- \psi)^n}$$  

$$\ell^{\ast}(\psi) = \ln(\binom{n}{v}) + v\ln(1 - \psi) - n\ln(2 - \psi)$$  
$$\ell^{\ast}(\psi) = \ln(\binom{170}{8}) + 8\ln(1 - \psi) - 170\ln(2 - \psi)$$  

$$\frac{d}{d\psi}\ell^{\ast}(\psi) = -\frac{8}{1 - \psi} + \frac{170}{2 - \psi} = 0$$  
$$8\psi - 16 - 170\psi + 170 = 0$$  
$$154 = 162\psi$$  
$$\widehat{\psi}_0^{mle} = \frac{154}{162} \approx 0.9506$$  

### Second Derivative test

We check the value of the second derivative at the MLE we found to verify it as a maximum:  

$$\frac{d^2}{d\psi^2}\ell^{\ast}(\psi) = -\frac{8}{(1 - \psi)^2} + \frac{170}{(2 - \psi)^2}$$  
$$\frac{-8}{(1 - 0.9506)^2}+\frac{170}{(2 - 0.9506)^2} \approx -3123.834... < 0$$  
Thus, the critical point we found is a maximum and verifies it as the MLE for $\psi$.  

## Bayesian Inference
Detail the Bayesian approach.

# Results
Present your findings. 

# Discussion / Conclusion
Discuss / conclude here.

# Bibliography

Brown, B. (2024). *Lecture Title*. Lecture slides, Course Name, University Name.

Doe, J. (2020). Title of the Paper. *Journal Name*, 12(3), 45-67.

Last, F., & Last, F. (2025). *Book Title*. Publisher.

Smith, A., & Johnson, C. (2023). *Title of the Online Article*. Retrieved from https://www.example.com.

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

## Proofs
If applicable, include detailed mathematical derivations or additional theoretical explanations.
